### 2022-11-09 NSF2022/Description/specificaim2.tex
first draft of Newton solver sec

% ----------------------------------------------------------------------
\subsubsection{Iterative methods for the Allen-Cahn HAP model}
\label{subsec:AC}

\begin{wrapfigure}[28]{l}{1.8in}
  \vspace{-8pt}
  \centering
  \input{figures/doublewell.tikz} 
  \caption{\label{fig:CA} \footnotesize (a) A double well potential. (b)
  The solution of the one-dimensional phase
  transition~\eqref{eqn:phase}. (c) The convergence of exact and inexact
  Newton's methods.}
\end{wrapfigure}
The PIs have developed numerical methods to solve~\eqref{eqn:phase} when
$W(\phi) = \phi^2$---this choice results in the linear screened Laplace
equation. For a general double well potential $W$, with local minima at
$\phi_0$ and $\phi_1$, we will interpret the solution
of~\eqref{eqn:phase} as the root finding problem for
\begin{align}
  \label{eqn:phaseRoot}
  J[\phi] = -\epsilon^2 \Delta \phi + \tfrac{1}{2}W'(\phi).
\end{align}
Equation~\eqref{eqn:phaseRoot} will be solved with Newton's method,
$\phi^{N+1} = \phi^{N} - \alpha (dJ^N)^{-1} J(\phi^N)$, where $dJ^N =
-\epsilon^2 \Delta + \tfrac{1}{2}W''(\phi^N)$ is the Fr\'{e}chet
derivative of $F$, and $\alpha$ is the step size that will be determined
using backtracking. As expected, the biggest computational challenge is
solving $-\epsilon^2 \Delta v + \tfrac{1}{2}W''(\phi^{N}) v =
J[\phi^N]$. Solving this variable coefficient PDE with an integral
equation is challenging since a fundamental solution is not readily
available. However, by choosing a double well potential with
$W''(\phi_0) = W''(\phi_1) = \beta$, we propose to instead use an
inexact Newton's method where $dJ^{-1}$ is approximated by solving
$-\epsilon^2 \Delta v + \tfrac{\beta}{2} v = J[\phi^N]$. We test the
efficacy of exact and inexact Newton's method applied
to~\eqref{eqn:phaseRoot} in $(0,1)$ by discretizing $dJ$ with a
second-order finite difference method. The double well potential is
illustrated in Figure~\ref{fig:CA}(a), and both the exact and inexact
Newton's methods converge to the solution in Figure~\ref{fig:CA}(b).
Figure~\ref{fig:CA}(c) shows the max norm of the residual at each Newton
iterate when using the exact second derivative (red curve) and the
approximate second derivative (blue curve). While the inexact Newton's
method requires roughly twice as many iterations to reach of a tolerance
of $10^{-4}$, each iteration is significantly cheaper to apply.

Having demonstrated that the inexact Newton method converges in one
dimension, it will be coupled with IE formulations to
solve~\eqref{eqn:phase} in two and three dimensions. The solution of
$-\epsilon^2 \Delta v + \tfrac{\beta}{2} v = J[\phi^N]$ can be written
as $v = \phi^P + \phi^H$ where $\phi^P$ is a volume potential and
$\phi^H$ is a layer potential. The volume potential is
\begin{align*}
  \phi^{P}(\xx) = \int_{\Omega(t)} G(\xx,\yy) J[\phi^N](\yy)\, d\yy, 
    \quad \xx \in \Omega(t),
\end{align*}
where $G$ is the fundamental solution of $-\epsilon^2 \Delta +
\tfrac{\beta}{2}$. Volume potentials are more challenging to compute
numerically than layer potentials. However, very recently, a suite of
new methods have been developed to accurately and efficiently compute
volume potentials~\cite{fry-leh-tor2018, and-zhu-vee2022}. The boundary
condition will be enforced using $\phi^H(\xx)$ which will be represented
as a double-layer potential. This approach of combining volume and layer
potentials has successfully been used by PI BQ~\cite{kro-qua2010} and
others~\cite{fry-kro-tor2019, kli-ask-kro2020, che-hua-lei2005}.

>>>>

% ----------------------------------------------------------------------
\subsubsection{An inexact Newton solver for the phase transition PDE}
\label{subsec:AC}

\begin{wrapfigure}[28]{l}{1.8in}
  \vspace{-8pt}
  \centering
  \input{figures/doublewell.tikz} 
  \caption{\label{fig:CA} \footnotesize (a) A double well potential. (b)
  The solution of the one-dimensional phase
  transition~\eqref{eqn:phase}. (c) The convergence of exact and inexact
  Newton's methods.}
\end{wrapfigure}
The PIs have developed numerical methods to solve~\eqref{eqn:phase} when
$W(\phi) = \phi^2$---this choice results in the linear screened Laplace
equation. For a general double well potential $W$ with local minima at
$\phi_0$ and $\phi_1$, we propose to interpret the solution as the root
finding problem for
\begin{align}
  \label{eqn:phaseRoot}
  J[\phi] = -\epsilon^2 \Delta \phi + \tfrac{1}{2}W'(\phi).
\end{align}
Equation~\eqref{eqn:phaseRoot} will be solved with Newton's method,
$\phi^{N+1} = \phi^{N} - \alpha (dJ^N)^{-1} J(\phi^N)$, where $dJ^N =
-\epsilon^2 \Delta + \tfrac{1}{2}W''(\phi^N)$ is the Fr\'{e}chet
derivative of $F$, and $\alpha$ is the step size that will be determined
using backtracking. This requires Solving the variable coefficient PDE
$-\epsilon^2 \Delta v + \tfrac{1}{2}W''(\phi^{N}) v = J[\phi^N]$ with an
integral equation, but this is challenging since a fundamental solution
is not readily available. However, by choosing a double well potential
with $W''(\phi_0) = W''(\phi_1) = \beta$, we propose to instead use an
inexact Newton's method where $dJ^{-1}$ is approximated by solving
$-\epsilon^2 \Delta v + \tfrac{\beta}{2} v = J[\phi^N]$. We test the
efficacy of the exact and inexact Newton's method applied
to~\eqref{eqn:phaseRoot} in $(0,1) \subset \RR$ by discretizing $dJ$
with a second-order finite difference stencil. The double well potential
is illustrated in Figure~\ref{fig:CA}(a), and both the exact and inexact
Newton's methods converge to the solution in Figure~\ref{fig:CA}(b).
Figure~\ref{fig:CA}(c) shows the norm of the residual at each Newton
iterate when using the exact second derivative (red curve) and the
approximate second derivative (blue curve). While the inexact Newton's
method requires roughly twice as many iterations to reach of a tolerance
of $10^{-4}$, each iteration is significantly cheaper to apply.

Having demonstrated that the inexact Newton method converges in one
dimension, it will be coupled with IE formulations to
solve~\eqref{eqn:phase} in two and three dimensions. The solution of
$-\epsilon^2 \Delta v + \tfrac{\beta}{2} v = J[\phi^N]$ can be written
as $v = v^P + v^H$ where $v^P$ is a volume potential and
$v^H$ is the double layer potential. The volume potential is
\begin{align*}
  v^{P}(\xx) = \int_{\Omega(t)} G(\xx,\yy) J[\phi^N](\yy)\, d\yy, 
    \quad \xx \in \Omega(t),
\end{align*}
where $G$ is the fundamental solution of $-\epsilon^2 \Delta +
\tfrac{\beta}{2}$. Volume potentials are more challenging to numerically
compute than layer potentials, but methods have recently been developed
to accurately and efficiently compute volume
potentials~\cite{fry-leh-tor2018, and-zhu-vee2022}. The boundary
condition will be enforced using $\phi^H(\xx)$ which will be represented
as a double-layer potential. This approach of combining volume and layer
potentials has successfully been used by PI BQ~\cite{kro-qua2010} and
others~\cite{fry-kro-tor2019, kli-ask-kro2020, che-hua-lei2005}.
